---
title: "Data Science for Public Policy"
subtitle: "Assingment 04"
author: "Nicholas Coukoulis, Moeko Kondo"
format: 
  html:
    code-line-numbers: true
execute: 
  warning: false
urlcolor: blue
editor_options: 
  chunk_output_type: console
---

# Question
In the United States, the shortage of public school teachers is critical problem over the years.
How serious the problem is? What is the etiology of the problem? 

```{r}

```

# graph 4
Data Source: [National Center for Education Statistics, Digest of Education Statistics 2021](https://nces.ed.gov/programs/digest/d21/tables/dt21_228.12.asp)

```{r}
library(readxl)
library(ggplot2)
library(tidyr)
library(tidyverse)

gunshooting <- read_excel("/Users/momom/Desktop/introdata/groupproject/Assignment04/gunshooting.xls", 3)
View (gunshooting)

df <- pivot_longer(data = gunshooting, cols=c(Deaths, Injuries), names_to="variable", values_to="value")

ggplot(data = df, mapping = aes(x = year, y = value, fill = variable)) +
  geom_area() +
  scale_fill_manual(values = c("#9966CC", "#E1BEE7")) +
  scale_x_continuous(
    expand = expand_scale(mult=c(0, 0.01)),
    breaks = unique(df$year)) +
  scale_y_continuous(
    limits = c(0, 185),
    breaks = seq(0, 185, by = 30),
    expand = c(0, 0),
    position = "left") +
  labs(
  title = paste("School Gun Violence"),
  subtitle = "Increase trend in the number of gun shooting deaths and injuries in elementary and secondary \nschools from 2000 to 2020. ",
  caption = "Data: National Center for Education Statistics, Digest of Education Statistics 2021",
  x = "", y ="") +
  theme_minimal() +
  theme(panel.border = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.caption = element_text(hjust = 0),
        legend.position = "top",
        legend.justification = "left",
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 30))

```

Gun violence has surged in recent years, with 46 deaths and 72 victims of cancer violence in Elementary and Middle Schools in the 2020-2021 school year despite the Covid19 shutdown. Teachers and school staff responsible for protecting students in schools are taking an increasingly heavy physical and emotional toll; [another study by the National Center for Education](https://nces.ed.gov/programs/digest/d21/tables/dt21_229.20.asp) found that reported incidents, including violent incidents and thefts, totaled 29.2 per 1000 students (2019-2020 school year). The safety of the working environment for teachers is threatened and is one of the major factors contributing to the teacher shortage. 


## Stretch Exercise 
### Part 1: .gitignore 
When adding the data file added included in the gitignore file, the following message was returned: 
The following paths are ignored by one of your .gitignore files:
teachersalaries_neaestimates_2021.xlsx
hint: Use -f if you really want to add them.
hint: Turn this message off by running
hint: "git config advice.addIgnoredFile false"


