---
title: "Data Science for Public Policy"
subtitle: "Assignment 04"
author: "Nick Coukoulis, Moeko Kondo"
format: 
  html:
    code-line-numbers: true
    code-fold: true
execute: 
  warning: false
urlcolor: blue
editor_options: 
  chunk_output_type: console
---
## Graph 1
### The Decline of Teaching as a Worthwhile Profession in the U.S. 
```{r}
# Teacher Careerpath Attitudes Graph
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyr)

teacherattitude <- read_excel("/Users/nicholascoukoulis/Documents/Georgetown/MPP/Spring2023/DataScience/assignments/assignment04/Assignment04/Teacher as Career_PDK Poll.xlsx")
View (teacherattitude)

teacherattitude_long <- teacherattitude %>% 
  pivot_longer(
    cols = 2:3,
    names_to = "viability",
    values_to = "responses",
    values_drop_na = TRUE
)

teacherattitude_long$Date <- as.Date(teacherattitude_long$Date, format = "%Y-%m-%d")
lim = as.Date(c("1969-01-01", "2022-30-11"))

ggplot(data = teacherattitude_long, mapping = 
         aes(x = `Date`, 
             y = responses, 
             group = 1)) +
  geom_line(data = teacherattitude_long, aes(group = viability, colour = viability)) +
  labs (title = (paste("The Decline of Teaching as a Worthwhile Profession in the U.S.")),
        subtitle = (paste ("'Would you like to have a child of yours take up teaching in public schools as a career?'")),
        caption = "Data: Annual Phi Delta Kappa Poll; selected years, 1969 through 2022.",
        x= "",
        y = "") +
  coord_cartesian(ylim = c(0,100))+
  scale_x_date(limits = lim,
               date_breaks = ("3 years"),
               date_labels=("%Y"))+
  scale_y_continuous (breaks = c(25, 50, 75, 100))+
  scale_color_manual(name = "United States, % responding 'yes' or 'no'",
                     values = c("No" = "#FF0000", "Yes" = "#1879CE"))+
  guides(color = guide_legend(override.aes = list(size = 3) ) )+
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "top",
        legend.justification = "left",
        plot.caption = element_text(hjust = 0))
```

## Graph 2
### U.S. Public K-12 Education Job Openings vs. Hires
```{r}
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyr)
library(zoo)

openingshires <- read_excel("/Users/nicholascoukoulis/Documents/Georgetown/MPP/Spring2023/DataScience/assignments/assignment04/Assignment04/jobopenings_hires_bls_3.2.23.xlsx")
View (openingshires)

openingshires_long <- openingshires %>% 
  pivot_longer(
    cols = 3:14,
    names_to = "month",
    values_to = "value",
    values_drop_na = TRUE,
    values_transform = list(value=as.character),
  )

openingshires_long <- openingshires_long %>% 
  mutate(date = as.yearmon(paste(month, year)))  %>%
  group_by(date)

openingshires_long$value <- as.numeric(as.character(openingshires_long$value))

lim = as.yearmon(c("Jan 2012", "Dec 2022"))

ggplot(data = openingshires_long, mapping = 
         aes(x = `date`, 
             y = value,
             group = status)) +
  geom_rect(aes(
    xmin = as.yearmon("Mar 2020"),
    xmax = as.yearmon("Dec 2022"),
    ymin = -Inf,
    ymax = Inf),
    alpha =0.1,
    fill = "#E9EDF0")+
  geom_line(data = openingshires_long, aes(group = status, colour = status))+
  scale_x_yearmon(limits = lim,
                  expand = c(0, 0),
                  breaks = unique(openingshires_long$year),
                  format=("%Y"))+
  scale_y_continuous (breaks = c(50, 100, 150, 200)) +
  coord_cartesian(ylim = c(0,220))+
  labs (title = (paste("Job Openings in K-12 Public Education Eclipse Hires Nationally")),
        subtitle = (paste ("A troubling gap began to emerge before the pandemic that has only been compounded.")),
        caption = "Data: U.S. Bureau of Labor Statistics, Job Openings and Labor Turnover Survey (seasonally adjusted data)",
        x= "",
        y = "Amount in Thousands") +
  scale_color_manual(name = "",
                     values = c("openings" = "#FF0000", "hires" = "#758D99"), labels= c("Hires", "Openings"))+
  guides(color = guide_legend(override.aes = list(size = 3) ) )+
  theme_minimal() + 
  theme(panel.border = element_blank(), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "top",
        legend.justification = "left",
        plot.caption = element_text(hjust = 0))

```

## Graph 3
### In Which States do Teachers Earn the Most? 
```{r}
#Lollipop Salaries Graph
library (ggplot2)
library(dplyr)
library(readxl)
library(forcats)
library(tidyr)
library(stringr)

teachersalary <- read_excel("/Users/nicholascoukoulis/Documents/Georgetown/MPP/Spring2023/DataScience/assignments/assignment04/Assignment04/teachersalaries_neaestimates_2022.xlsx")
View (teachersalary)

teachersalary <- teachersalary %>%
  arrange(salary) %>%
  mutate(state = factor(state, levels = .$state))

ggplot(teachersalary, aes(salary, state, color = "#E3120B"))  +
  geom_segment(aes(x = 0, y=state, xend = salary, yend = state), color = "grey50") +
  geom_point()+
  geom_vline(xintercept = 67507, linetype = "dotted") +
  labs (title = (paste("In Which States do Teachers Earn the Most?")),
        subtitle = (paste ("Average U.S. K-12 Public Teacher Salary Estimates by State")),
        caption = "Data: National Education Association Rankings and Estimates Report, 2022",
        x= "",
        y = "") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(), 
        legend.position = "none",
        plot.caption = element_text(hjust = 0))
```

## Stretch Exercise 
### Part 1: .gitignore 
When adding the data file added included in the gitignore file, the following message was returned: 
  The following paths are ignored by one of your .gitignore files:
  teachersalaries_neaestimates_2021.xlsx
  hint: Use -f if you really want to add them.
  hint: Turn this message off by running
  hint: "git config advice.addIgnoredFile false"
  



