---
title: "Data Science for Public Policy"
subtitle: "Assingment 04"
author: "Nicholas Coukoulis, Moeko Kondo"
format: 
  html:
    code-line-numbers: true
execute: 
  warning: false
urlcolor: blue
editor_options: 
  chunk_output_type: console
---

# Question
In the United States, the shortage of public school teachers is critical problem over the years.
How serious the problem is? What is the etiology of the problem? 

```{r}

```

# graph 4
Data Source: [National Center for Education Statistics, Digest of Education Statistics 2021](https://nces.ed.gov/programs/digest/d21/tables/dt21_228.12.asp)

```{r}
library(readxl)
library(ggplot2)
library(tidyr)
library(tidyverse)

gunshooting <- read_excel("/Users/momom/Desktop/introdata/groupproject/Assignment04/gunshooting.xls", 3)
View (gunshooting)

df <- pivot_longer(data = gunshooting, cols=c(Deaths, Injuries), names_to="variable", values_to="value")

ggplot(data = df, mapping = aes(x = year, y = value, fill = variable)) +
  geom_area() +
  scale_fill_manual(values = c("#9966CC", "#E1BEE7")) +
  scale_x_continuous(
    expand = expand_scale(mult=c(0, 0.01)),
    breaks = unique(df$year)) +
  scale_y_continuous(
    limits = c(0, 185),
    breaks = seq(0, 185, by = 30),
    expand = c(0, 0),
    position = "left") +
  labs(
  title = paste("School Gun Violence"),
  subtitle = "Increase trend in the number of gun shooting deaths and injuries in elementary and secondary \nschools from 2000 to 2020. ",
  caption = "Data: National Center for Education Statistics, Digest of Education Statistics 2021",
  x = "", y ="") +
  theme_minimal() +
  theme(panel.border = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.caption = element_text(hjust = 0),
        legend.position = "top",
        legend.justification = "left",
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 30))

```

Gun violence has surged in recent years, with 46 deaths and 72 victims in elementary and secondary Schools in the 2020-2021 school year despite the shutdown during the Covid-19 pandemic. Teachers and school staff responsible for students' safety in schools bear an increasingly heavy physical and psychological burdens. In addition, [another study by the National Center for Education](https://nces.ed.gov/programs/digest/d21/tables/dt21_229.20.asp) found that reported incidents, including not only gun violence but other violent incidents and thefts, totaled 29.2 per 1000 students (2019-2020 school year). Teachers are facing a situation where they have to work on their correctional education as well as secure their student safety, while their own safety is threatened.
As such, serious school violence is one of the major factors contributing to the teacher shortage. 


## Stretch Exercise 
### Part 1: .gitignore 
When adding the data file added included in the gitignore file, the following message was returned: 
The following paths are ignored by one of your .gitignore files:
teachersalaries_neaestimates_2021.xlsx
hint: Use -f if you really want to add them.
hint: Turn this message off by running
hint: "git config advice.addIgnoredFile false"


