---
title: "Data Science for Public Policy"
subtitle: "Assingment 04"
author: "Nicholas Coukoulis(), Moeko Kondo(830159968)"
format: 
  html:
    code-line-numbers: true
execute: 
  warning: false
urlcolor: blue
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)

glimpse(storms)

ggplot(storms) +
  geom_point(mapping = aes(x=wind, y=pressure), color="red")


```

#test message

# Question
```{r}
# In the US, the shortage of public school teachers is critical problem over the years.
# How serious the problem is? What is the etiology of the problem? 
```


```{r}

```

# graph 4
```{r}
library(readxl)
library(ggplot2)
library(tidyr)

gunshooting <- read_excel("/Users/momom/Desktop/introdata/groupproject/Assignment04/gunshooting.xls", 3)
View (gunshooting)

df <- pivot_longer(gunshooting, cols=c(Deaths, Injuries), names_to="variable", values_to="value")

ggplot(data = df) +
  geom_bar(mapping = aes(x = year, y = value, fill = variable),stat="identity", position="dodge")

```

## Stretch Exercise 
### Part 1: .gitignore 
When adding the data file added included in the gitignore file, the following message was returned: 
The following paths are ignored by one of your .gitignore files:
teachersalaries_neaestimates_2021.xlsx
hint: Use -f if you really want to add them.
hint: Turn this message off by running
hint: "git config advice.addIgnoredFile false"


